Asymmetric relationships and their effects on coexistence: Supporting information

### 2023-07-17 paper/RefereeResponses1.tex

\noindent ***\emph{We address your questions and comments below.} \\
%DONE

\noindent C4) My biggest question is related to the removal of ATAs from a given true bivariate distribution.  This is described in SI Section S2 and shown visually for an example in SI Figure S1.

My first comment related to this is that some form of this text (whcih is relatively short---around half a page in the SI) and, importantly, the Figure should be in the main text.  I think this is such a key part of the definitions that are being used for the rest of the paper that readers will want to get an intuition for what the "partial sharp" distributions look like and how they are arrived at. \\

\noindent ***\emph{We have done what the referee requests, here, and we agree with this request. Thanks for
making the request. This request is somewhat in tension with other requests, from other referees, 
to minimize the quantitative 
content of the main text, and to keep the overall length under the journal limit while adding a bunch of
verbal explanations requested by the other referees. We point out this tension here, and refer back to it in
other responses, below. We also point out that we have not only moved the figure back to the main text,
we have also added a panel to the figure in response to the related comment C13, below.} \\
%DONE
%DAN: In other replies, you can refer to this reply to justify when you
%decide to add explanations to the SI instead of to the main text, and when you have to respond to requests
%to minimize the quant.

@@

\noindent C5) Perhaps equally important to clarifying this process for the readership is whether this process is unique. I could not quite get a sense of whether there could be alternative versions of this removal of ATA from a bivariate distribution, that still satisfy the authors' constraints on marginal distributions and overall correlation between the two variables. Are there alternatives, or is the authors' procedure unique in achieving the desired outcome?  If the former, then I would want to understand if the outcomes downstream tend to be robust to the specific method used. If the latter, it would be helpful to have an explanation as to why this is a unique decomposition. \\

\noindent ***\emph{This is a good question, thanks for asking it. The extent to which the process is 
unique is now covered in a new SI section, refered to from the main text at the end of section 2.3, 
right at the end of the text the referee asked us to move to the main text. There is a choice involved,
but it is pretty canonical -- certainly more canonical than the kinds of choices which are often made
in MCT, e.g., $E$ and $C$ can sometimes be defined in more than one way. Basically the choice involves 
using a bivariate normal distribution instead of another, more esoteric 
distribution with standard normal marginals and symmetric tail association. The bivariate normal is by far the 
most natural example of such a distribution. See the new SI section for further explanations. }

\emph{We are reticent, at this time, to launch into a lot of robustness explorations with regard to this choice, 
as the referee may be suggesting, for several reasons. First, other referees and the editor have urged us to
simplify, and such explorations would complexify instead. Second, as mentioned above, the choice we made is very
natural and quite standard, so exploring other possible choices seems therefore of 
secondary importance. Finally, our overall goal in writing this paper
was to show that ATAs may be important for coexistence and, as such, deserve additional exploration in this
context by the wider research community. As long as any choices we made are reasonable (they are), our point is
made, regardless of whether making some more esoteric choices might have altered outcomes. As we already argue in the
Discussion, since we have now showed that ATAs \textbf{can} be important, future work should explore \textbf{when} 
they are important, and part of 
that exploration could be whether some more esoteric choice for $(e_i c_{i \bs i})$ might alter our results. }\\
%DONE

@@

\noindent C6) The notation by which the partial sharp variables are introduced was a little unintuitive to me.  For example, when the authors compare the expectation values 
$E[r(E,C)]$ and $E[r(E^{||},C^{||})]$, what is really changing is the distribution over which the expectation values are evaluated.  Maybe this is standard notation, but somehow it may be clearer to put the subscript on the E for expectation value, to indicate that the probability distribution used to generate the definition of the averaging is what differs between the two expressions, while the variables E and C themselves are in the end just dummy variables to be integrated over.  \\

\noindent ***\emph{This comment has led us to put some clarifying technical comments in a new SI section 2,
so thanks for making the comment.}

\emph{We use notational conventions of random variables which are formally correct and standard, 
as we now explain in more detail in the SI. 
$E_i$, $C_{i \bs i}$, $E_i^{\nat}$, $C_{i \bs i}^\nat$ and the earlier 
notation $E_i^\shp$ and $C_{i \bs i}^\shp$ all denote \textbf{random variables}, using a 
notational approach which is standard in applications of probability. 
Whereas the \textbf{distributions} of the random variables $E_i$, $E_i^{\nat}$ and $E_i^\shp$ (for instance) are all the same,
these things are \textbf{not} the same \textbf{as random variables} because their relationships with other random variables 
differ. The relationships between random variables are considered instrinsic
to the notation. Thus it's not that somehow the expected value in $\E[r(E,C)]$ is a different definition of expected
value than the one in $\E[r(E^{||},C^{||})]$ -- there is only one expected value operator. 
Instead, $r(E,C)$, $r(E^{||},C^{||})$ and $r(E^\shp,C^\shp)$ differ from each other 
as random variables because of the distinct relationships between their
different $E$ and $C$ constituents. It would be formally 
incorrect, and confusing to readers familiar with probability, to try to define multiple expected value 
operators, as the referee seems to suggest, when what fundamentally differs, here, is the random variables 
themselves. Readers familiar with
probability already know what the one accepted definition of expected value is, would typically
not be comfortable with providing additional definitions, and are comfortable handling these concepts
with the notational conventions we have used. To remind the referee of some concepts from basic probability
which illustrate these conventions, recall that if $X$ and $Y$ are random variables, the expectation of the 
product, $\E(XY)$, depends not only on the univariate distributions of $X$ and $Y$, but also on the relationship
between $X$ and $Y$, namely their joint distribution. In particular, $\E(XY)=\E(X)\E(Y)$ if $X$ and 
$Y$ are independent, but not necessarily otherwise. These definitons and notation are completely 
standard in probability theory; notice that we use the same definition of $\E$ regardless of 
whether $X$ and $Y$ are independent or nonindependent, it's just that the product random variable
$XY$ differs accordingly so you get different results of $\E(XY)$ in those distinct cases.}

\emph{Our notation also parallels that of Ellner et al (2016, 2019). To make it
easier for readers to understand and to reduce the potential for future confusion, it makes sense to continue
using existing notation whenever that notation is formally correct and otherwise effective (it is, as explained above).}

\emph{Nevertheless, the referee's comment has made us understand that some readers unfamiliar with random variable 
notation may benefit from putting some
of these explanations in the SI, so we have done so.} \\
%DONE

@@

\noindent C7) It was not clear to me what the reasoning was for including the correlation and ATAs in abundance for the two planktonic species in Figure 1(d) and (e).  These ATAs in abundances don't in an obvious (to me) way translate into ATAs of the underlying environment and competition variables for each species. Is this really a demonstration that ATAs occur for C and E?  If not, then it seems to be like saying ATAs exist somewhere in nature, and therefore might be present in our focal situation, which is not such a very strong reason to show (d) and (e).  Maybe I am missing something. \\

\noindent ***\emph{Yes, the referee is right about this, we are not illustrating ATAs in $E$ and $C$ at this stage of the 
paper - the reader would anyway not be equiped to handle that at this stage, since we have not even formally defined 
$E$ and $C$ at the point this figure is first cited (in the very first paragraph). 
We were asked to include this kind of empirical illustration 
by poeple who read the manuscript before we submitted it, so it seems important to some readers to have this. 
The value is not only in showing that ATAs occur in nature, but is priarily  
in helping people understand more fully what they look like when they occur by giving real examples. 
We appreciate what the referee is saying,
here, that perhaps one non-$EC$ example of ATAs from nature is not going
to give much motivation for looking for ATAs in $E$ and $C$. However, we have published droves of other examples of
ATAs in nature, and we now cite these (changing what used to be ``Fig. 1d,e show contrasting examples of ATAs in nature
using plankton population density time series'' to ``Fig. 1d,e show contrasting examples of ATAs in nature
using plankton population density time series; and many other examples have also been previously explored (Ghosh et al 2020)''). 
And anyway we cannot produce a prior example of ATAs between $E$ and $C$ because there are no prior examples - this paper 
is, to our
knowledge, the first time anyone has looked at ATAs between $E$ and $C$. Fortunately, we think the other motivational
content of the Introduction is strong, and should stand on its own to motivate readers to spend their mental 
energy reading this paper. In partcular, we show in the Introduction, for one simple simulation, that ATAs between E and
C make a visually striking difference between simulation outcomes that appears clearly to influence coexistence.} \\
%DONE

@@

\noindent ***\emph{We thank the referee for this viewpoint. We also thank the referee for acknowledging that this
is their opinion, and that data are insufficient for a clear conclusion to be reached at this point. We find that 
such an acknowledgment is both rare from a referee and very useful for causing authors to most usefully revise their 
paper. We have added new discussion material covering these topics and find that the new material improves the paper.
We tried to represent both viewpoints in the new material, and we cited the papers the referee mentions.}

\emph{We point out that we worked not only
with the lottery model, but also with an empirical diatom system. That system definitely supported the importance of 
ATAs for coexistance, which would tend to give at least one data point in favor of the importance of these phenomena, though we 
acknowledge that our lab systems is also a system ``where niche differentiation via ... predators is impossible.''
It may also be a system ``where niche differentiation via different resources ... is impossible,'' though resource 
differentiation can happen when you don't expect it. It seems to us 
the first step in assessing the importance of a newly recognized phenomenon has to be examining it in simple systems
such as we have used, prior to examing it in field systems, in all their complexity, where niche differentiation
by resources and predators can more readily occur. }

\emph{We also point
out (and the referee already seems to accept this) that it is possible, \emph{a priori}, for ATAs to be 
important for coexistence in a system even while storage effects are not.
Storage effects are ATA effects plus what we call the effects of $EC$ correlation \emph{per se} 
($\Delta_i^{[E \nat C]} in our notation), and either of these two summands can be negative or positive.
So if the twe effects are oppostite in sign and nearly equal, storage effects will be unimportant but ATA
effects will be important for coexistence.}

\emph{So the referee appears to agree with us that storage effects and ATA effects may or may not turn
out to be broadly important in ecology. Our research goals were to introduce the idea of ATA effects on
coexistence and to show that ATAs \text{may} be broadly important for coexistence, by showing examples where they
\textbf{are} important. The few examples we showed are all that are readily available at this time. 
Future work can assess, broadly, how important ATAs are by looking across a larger number of systems, once
available. These are two different goals. It seems crucial to us not to allow insufficiently tested 
opinions about the second set of goals to obstruct the publication of results pertinent to the first set of 
goals, so we thank the referee for not doing that.}

\emph{The new Discussion material is in a new SI section which is cited from the main text Discussion section. 
We would have liked to put some of the new material into the main text. But length limits, together with
clear requests from other referees to add a variety of content to the main text, made doing so impossible.}
%DAN: Write the discussion material which is promised and the citation from the Discussion.

% 
% Note also that Chu and Adler (2015) say "we now believe that the models we fit may not be appropriate for rigorously quantifying fluctuation-dependent mechanisms. For example, the temporal storage effect requires positive covariance between the effects of competition and the environment (Chesson and Huntly 1989): in years most favorable to a species, it should also be the most vulnerable to competition. However, we did not specify our vital rate regressions in a way that makes the covariation between the environment (or random year effects) and competition explicit. As a result, our model may not allow full expression of a temporal storage effect and the simulations may underestimate its strength. Such differences in model specification may explain why a previous analysis, based on a very different modeling approach, showed a strong storage effect in the Kansas community (Adler et al. 2006) in contrast to the weak effect we observed here. We plan to address this model specification issue in future work focused on the population and community consequences of environmental variability."
% 
% Note that the Stump and Vasseur paper very clearly says that they don't think storage effects are important. 
% It's a 73-page ms, I don't have time to absorb it. This will be part of acknowledging the viewpoint. Maybe Jasmin
% has time to read it and decide how it relates to our paper? Do the assumptions they explore also apply
% to the Ellner approach of quantifying storage effects?
} \\
%Dan to do

@@

\noindent C12) I like the conceptual idea of removing the long tail. \\

\noindent ***\emph{Thank you for the positive feedback.} \\

\noindent C13) However, I have somewhat difficulties to precisely understand/be sure I correctly understood what this does. I suggest you move a simpler version of Fig S1 into the main text and show the actual distribution, the uncorrelated distribution of $E$ and $C$, but same marginal distribution, and the distribution with the same covariance, but the ATA removed. Additionally, it might be beneficial to indicate which distribution is then used to compute which part of the decomposition. For example, in Figure S1 d you show $E_i^{||}$, but from my understanding this distribution would be used to compute $E_i^{(EC)}$, correct? \\

\noindent ***\emph{In comment C4 above, another referee asked us also to put S1 back into the main text, so
we have done as the referee asks here. The new Fig. 2 has the three elements the referee asks for, here, as 
its panels (a), (e) and (d). We retained the other panels as well, since those were requested by the 
other referee. Regarding the second part of the referee's comments, 
there is nothing in our paper for which the notation $E_i^{(EC)}$ is used, but we think what the referee must
be requesting here is what we remind the reader that $(E_i^\shp,C_{i \bs i}^\shp)$ is used to compute storage effects,
$\Delta_i^{(EC)}$, and that $(E_i^\nat,C_{i \bs i}^\nat)$ is used to compute ATA effects, $\Delta_i^{[EC]}$. We have 
now added such a reminder to the figure caption.} \\

@@

\noindent C29) L230 $\mu_1-\mu_2$ is negative, why not take $\mu_2-\mu_1$? \\

\noindent ***\emph{Well, we can do this if the editor insists, but to us, avoiding the use of
negative numbers is insufficient reason to have to undertake the substantial and time-consuming 
work (and potential for error) 
of re-doing several figures, and revising many formulas (in the main text and SI) which currently use the other convention.
After all, negative numbers are basic, and represent no real impediment to the average ecologist understanding the 
paper. And we could instead spend that time working on the concepts underlying the next paper.} \\
%Jasmin to consider (don't do yet) how hard/error prone this change would be, taking into account the 
%formulas in the SI? Unless this is MUCH easier and MUCH less error prone than Dan's hunch, I suggest 
%we don't do it.

\noindent C30) L316: So why do you show values of $\sigma >6$? And also maybe highlight this in the figure Figure 1d and e: I’m semi convinced by these distributions showing ATA. In 1d it simply appears that the variance scales with the density. 1e is quite unconvincing to me. That’s generally fine, but figure 1 seems to be a conceptual figure, so I wonder whether you don’t have a better example? \\

\noindent ***\emph{Regarding the first referee question listed above, well, what we are 
trying to do here is accurately portray our results about
when ATAs were important and when they were not. To us, the most forthright way of doing that is to
show \textbf{all} the results, including the values of $\sigma$ which are probably unrealistic, and mention
(as we do) that some values are unrealistic. The point of this paragraph is to outline the circumstances
we have explored, using our lottery-model approaches, for which ATAs were important or not, and including all
our results seems important when such is our goal. Furthermore, there's no real cost to showing
all the $\sigma$'s we explored, it does not take up more manuscript space to do so.
Some readers may also take the lesson, for future work, from these explorations that one of the things that probably 
matters for continuing to explore when ATAs are important and when not is probably the strength
of the noise impacting the system. Finally, whether a value of $\sigma$ is ``unrealistic'' is 
a bit of judgement call rather than a formal cutoff, to we hesitate to censor results 
based on such a thing.}

\emph{Regarding the second part of the referee comment, pertaining to Figure 1d and e, it is perplexing to us that 
this is mentioned here,
as the paragraph at L316 is not directly related to Fig. 1d-e. That figure
shows motivational empirical examples, distinct from the lottery model work discussed around L316. Nevertheless, we take
seriously the referee's comment that they are are not convinced there are ATAs in Fig. 1d-e, and ...
%Jasmin, let's discuss what to do here. We could work back to what specific result this is in 
%the original paper and extract from p-value info, or we could come up with a better example,
%or something else, not sure, let's discuss. Maybe also augment the figure to make it clearer?
%How about if Jasmin: 1) adds statistical support to caption; 2) adds to the figure some means of
%hopefully drawing people's attention to the right features of this thing.
} \\
